Topic models

### 2021-04-06 manuscript.Rmd
finalize revisions

---
title             : "Mapping Correlates of Adolescent Emotion Dysregulation: A Text-mining Systematic Review"
shorttitle        : "ADOLESCENT EMOTION REGULATION"

@@

Adolescence is a developmentally sensitive period for emotion regulation [@zimmermannEmotionRegulationEarly2014]. 
<!--Adolescents experience biological, cognitive, and social changes that prompt new emotional
experiences and tax regulatory abilities [@steinbergAgeOpportunityLessons2014].
These coalescing changes temporarily restrict adolescents' capacity for emotion regulation 
[@zimmermannEmotionRegulationEarly2014],
resulting in more frequent and intense (negative) emotions [see @silkAdolescentsEmotionRegulation2003].-->
Although most adolescents successfully develop mature emotion regulation abilities,
as many as one in five develop psychopathology [@leeAdolescentMentalHealth2014]
with roots in emotion regulation difficulties
[@aldaoEmotionregulationStrategiesPsychopathology2010].
Given the prevalence of emotion regulation difficulties in adolescence,
their implications for mental health<!--and social functioning [@braetEmotionRegulationChildren2014]; @reindlSocializationEmotionRegulation2016]-->
and downstream cost to society,
it is important to have a comprehensive overview of the factors associated with adolescents' emotion regulation.
<!--Of particular interest are potential risk factors that render some youth susceptible to emotion regulation difficulties.-->
Despite the abundance of publications in this field,
it is difficult to obtain a comprehensive understanding of emotion regulation in adolescence [@riedigerEmotionRegulationAdolescence2014]:
Different (sub)disciplines have approached the topic in disparate ways [@riedigerEmotionRegulationAdolescence2014],
without consistent terminology [@bariolaChildAdolescentEmotion2011], conceptual frameworks [@stifterEmotionRegulation2019], or an overarching theoretical framework
[@bussTheoriesEmotionalDevelopment2019].
This has prompted calls for the integration of knowledge across research areas [@riedigerEmotionRegulationAdolescence2014] and consolidation of empirical research into overarching theory [e.g., @bussTheoriesEmotionalDevelopment2019].
Formal methods for theory construction state that the first step in this endeavor is to identify well-established relevant *empirical phenomena*,
defined as stable and general features of the world [@borsboomTheoryConstructionMethodology2020].
This study used a text mining systematic review to provide a comprehensive map of the empirical phenomena relevant to adolescent emotion regulation.

Despite recent interest in theory construction method,
few specific techniques exist for identifying relevant phenomena, aside from expert opinion [@borsboomTheoryConstructionMethodology2020].
This paper argues that systematic reviews are a suitable method for quantifying which phenomena experts consider to be relevant.
Traditional narrative reviews are limited, however, by small convenience samples,
confirmation bias,
and an undue emphasis on positive results [@littellEvidencebasedBiasedQuality2008].
This paper aims to overcome these limitations by adopting a novel method:
the text mining systematic review (TMSR).
Relative to narrative reviews, text mining can cover arbitrarily greater corpora,
and gleans insights from the literature by means of a more transparent and reproducible process.
<!--Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008],
the inherent "fairness" of text mining gives emerging themes a chance to come to the fore.-->
This approach assumes the frequency with which phenomena recur in the literature to be indicative of their relevance.
If it is additionally assumed that the frequency with which phenomena are studied together is indicative of the extent to which researchers believe them to be associated,
then the co-occurrence of phenomena within publications can be interpreted as a rudimentary nomological network: a mapping of relationships between theoretically relevant phenomena [see @alaviAligningTheoryMethodology2018].
This network can serve as a starting point for the formulation of a proto-theory,
which would additionally involve abductive reasoning: specifying putative mechanisms to explain associations between the phenomena [@borsboomTheoryConstructionMethodology2020].
By mapping the phenomena relevant to adolescent emotion regulation,
the present work thus lays the groundwork for future theory development.

@@

p <- quote({
  # Set margins to 0
  par(mar=c(0.5,0.5,0.5,0.5),
      mfrow=c(1,2))
  plot(g1, edge.curved = 0, layout=l1,
       vertex.label.family = "sans",
       vertex.label.cex = 0.7,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "A")
  legend(x=-1.2, y=1.1, names(cat_cols), pch=legend_shapes, col=cat_cols, pt.bg=cat_cols, pt.cex=1.4, cex=.7, bty="n", ncol=1)
  plot(g2, edge.curved = 0, layout=l2,
       vertex.label.family = "sans",
       vertex.label.cex = 0.7,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "B")
})
saveRDS(g1, "bl_g1.RData")
saveRDS(g2, "bl_g2.RData")
saveRDS(l1, "bl_l1.RData")
saveRDS(l2, "bl_l2.RData")
# Save files

@@

```{r figbaseline, warning=FALSE, eval = TRUE, results='asis', out.width = "100%", fig.cap="Phenomena relevant to adolescents' emotion regulation according to theory (A) and narrative reviews (B; transparent circles indicate constructs also present in the theory)."}
knitr::include_graphics("baseline_network.png") 
```
```{r allnetworks, warning=FALSE, eval = FALSE, results='asis', out.width = "100%", fig.cap="Correlates of emotion regulation difficulties according to theory (A), narrative reviews (B), and term co-occurrence in author keywords (C) and abstracts (D). Size of lines and circles represents frequency. Dashed lines represent links not involving dysregulation. Transparent circles indicate constructs also represented in the theory (panel A)."}

@@

Although it has been argued that there is a paucity of theories *specific to* adolescent emotion regulation [see @bussTheoriesEmotionalDevelopment2019],
many *relevant* theories are commonly invoked in empirical work.
In order to contextualize the results of the present systematic review
and assess to what extent the identified phenomena through text mining mirror or complement existing theory,
a brief review of the existing theoretical landscape is provided.
Others have published more detailed reviews of theories of emotional development [@bussTheoriesEmotionalDevelopment2019] and emotion regulation in adolescence [@riedigerEmotionRegulationAdolescence2014].
The left panel of Figure \@ref(fig:figbaseline) visually summarizes the phenomena discussed in the theoretical literature,
along with whether these were discussed as putative causes, outcomes, indicators, or protective factors in relation to emotion regulation.
This establishes a baseline for the original contributions of the text minning analysis.
<!-- 
Our text mining analysis of the phenomena empirical literature is best understood against a backdrop of -->

Two of the most general theories invoked to frame developmental research are the bioecological model [@bronfenbrennerBioecologicalModelHuman2007], and the transactional model [@sameroffUnifiedTheoryDevelopment2010].
The bioecological model describes how the `r keyw("environment")`
shapes individual development.
At the individual level, every person is imbued with `r keyw("biological")` predispositions,
and develops over time in interaction with contextual influences.
The most immediate source of contextual influences is the microsystem,
composed of people close to the individual.
Other influences stem from the macrosystem, consisting of `r keyw("political")` and `r keyw("economic")` influences,
and the exosystem, consisting of `r keyw("cultural")` norms and `r keyw("values")`.
The transactional model [@sameroffUnifiedTheoryDevelopment2010] is compatible with the bioecological model,
but places a stronger emphasis on development as a product of reciprocal influences between child and environment.
The transactional model distinguishes between proximal influences, which roughly correspond to the microsystem in the bioecological model, and distal influences, which derive from
structural factors indirectly shaping development, like `r keyw("socio-economic status")`,
`r keyw("schools")`, and the `r keyw("community")` (macro- and exosystem).
With increasing age, distal influences gain ground on proximal influences.
These two theories have a broader scope than most.
This means that they can be invoked to contextualize any developmental study,
but lack specificity - a shortcoming that curtails a theory's utility in generating hypotheses.
Domain-specific theories, by comparison, offer greater specificity.

One such domain-specific theory is Hall's notion of "storm and stress" [see
@arnettAdolescentStormStress1999].
It describes how `r keyw("hormonal")` changes diminish `r keyw("self-control")` and increase `r keyw("reactivity")`, which in turn leads to emotion
regulation difficulties, increased `r keyw("conflict with parents")`, and `r keyw("risky behavior")`.
In contemporary theory, this notion of diminished self-control and increased emotional reactivity is recast as a normative change that facilitates emotional maturation at the risk of emotional
disturbance [@arnettAdolescentStormStress1999] 
A potential limitation of conceptualizing emotion regulation difficulties as normative aspects of development is that adolescents' diverging destinies are underemphasized [cf. @croneUnderstandingAdolescencePeriod2012].
<!--SH I am still not sure what the empirical contribution is - can you be more specific with it in this sentence? -->
<!--Several recent publications have undertaken the
monumental task of providing a comprehensive overview of theories of emotional
development [e.g., @bussTheoriesEmotionalDevelopment2019;
@holodynskiDevelopmentEmotionsEmotion2006;
@riedigerEmotionRegulationAdolescence2014], and of the empirical work
pertaining to emotions in adolescence
[@coe-odessEmergentEmotionsAdolescence2019].  -->

One theory of normative emotional development [@sroufeEmotionalDevelopmentOrganization1995] focuses on developmental influences to a greater extent.
This theory posits that, as children grow older,
their increasing `r keyw("self-regulatory")` abilities
drive a transition from external emotion regulation by `r keyw("primary caregivers")`
toward autonomous emotion regulation.
This theory emphasizes `r keyw("social")` and `r keyw("cognitive influences")` as 
drivers of development. 
Social influences mainly occur
through `r keyw("parental co-regulation")`, `r keyw("parenting behaviors")`, and parent-child
`r keyw("attachment")`.
Cognitive influences occur through the development of the `r keyw("central
nervous system")` (CNS), `r keyw("cognition")`, and `r keyw("self-regulation")`.
<!--In terms of level of analysis, Sroufe's and Hall's theories corresponds approximately to the individual- and microsystem level of the ecological model [@bronfenbrennerBioecologicalModelHuman2007].-->
This theory's relevance for adolescent research is somewhat diminished by its focus on early childhood,
although there is substantial continuity in developmental influences throughout childhood and adolescence.
For instance, there is substantial similarity between mothers' and fathers' unique roles in emotional development in childhood [@vanlissaMothersFathersQuantitative2020] and adolescence [@vanlissaRoleFathersMothers2019].
The emphasis on socialization and `r keyw("neurocognitive development")`,
which is evident from both theories,
is the focus of several theories with a narrower scope.

One widely cited theory of socialization is the tripartite model [@morrisRoleFamilyContext2007].
It describes three pathways through which `r keyw("parents")` shape emotion regulation: `r keyw("modeling")`, `r keyw("parenting practices")`, and the
emotional `r keyw("family climate")`,
which subsumes attachment and `r keyw("marital relationship")` quality.
This theory acknowledges unique contributions by `r keyw("mothers")` and `r keyw("fathers")`, and the importance of siblings.
Others have adapted the tripartite model to describe peer socialization [@reindlSocializationEmotionRegulation2016], thus illustrating its wider usefulness.
A more abstract take on socialization is found in the internalization model of emotional development [@holodynskiDevelopmentEmotionsEmotion2006]. 
This theory describes the role of emotion in `r keyw("communication")`, and the `r keyw("cultural")` symbolic function of emotion.

In research on neurocognitive aspects of adolescents' emotion regulation,
`r keyw("polyvagal theory")` has been influential [@porgesOrientingDefensiveWorld1995].
This theory examines `r keyw("emotional experience")` and -regulation in relation to autonomous nervous system functioning, respiratory sinus arrhythmia, and the stress response.
Although polyvagal theory is relevant for development,
it does not explicitly address it.
Among developmental neurocognitive theories,
there is some consensus that the developmental asymmetry between `r keyw("motivational")`-emotional and inhibitory `r keyw("brain")` circuits gives rise to a "maturity gap" in middle adolescence [@croneUnderstandingAdolescencePeriod2012].
According to the model of social-affective engagement and `r keyw("goal flexibility")` [@croneUnderstandingAdolescencePeriod2012],
adolescents' cognitive engagement is dynamically responsive to
social and motivational `r keyw("goal salience")`.
This flexibility <!--, on the one hand,-->
prepares adolescents to <!--effectively engage cognitive systems in novel
challenging situations in a way that facilitates--> develop mature regulatory
abilities, but also places them at risk of impulsivity in
pursuit of `r keyw("peer approval")`.
To a greater extent than related writings [cf. @craccoEmotionRegulationChildhood2017], this model focuses on adolescents diverging destinies;
why some youngsters flourish while others languish [@croneUnderstandingAdolescencePeriod2012].
This is important for understanding individual differences in development.
Although the relevance of this theory is enhanced by its focus on adolescence,
it only tangentially addresses emotion regulation.
Furthermore, whereas this theory addresses cognitive factors and peers as drivers of development, it underemphasizes other relevant phenomena, such as parenting [cf. @morrisRoleFamilyContext2007].
Micro-scale theories like these are useful in explaining specific phenomena in detail,
but do not provide a comprehensive understanding of adolescent's emotion regulation.
To this end, it is important to consider them in the context of a larger framework of relevant phenomena.

<!--Like Sroufe, Holodynski describes an age graded transition from
interpersonal to intrapersonal emotion regulation.
In doing so, Holodynski applies Vygotsky's theory of development to the domain of
emotion, and presents a detailed integrated model of emotional experience and
regulation.-->
Theories of the phenomenon of emotion regulation offer insight into
intra-individual drivers of emotion regulation development.
The influential process model [@grossHandbookEmotionRegulation2013] describes the
phenomenon of emotion regulation, from eliciting cue to ultimate response.
This model posits that individuals use `r keyw("strategies")` to modulate the different stages of this process,
consciously or otherwise. 
Individuals who engage in maladaptive emotion regulation strategies tend to experience more negative emotions,
diminished well-being, and greater strain in interpersonal relationships
[ @bellRelationshipsInputsOutputs2000].
However, comparative studies indicate that the adaptive versus maladaptive psychosocial consequences of specific strategies appear to be partly contingent on the `r keyw("cultural context")` [see
@bariolaChildAdolescentEmotion2011].
Similar to the process model, the "social
information processing" theory also describes the role of cognitive
processes and strategies in emotion regulation
[@lemeriseIntegratedModelEmotion2000].
One shortcoming of these theories for understanding adolescents' emotion regulation 
is the lack of a developmental component.

@@

Despite the abundance of theory *relevant* to emotion regulation in adolescence,
the literature has several limitations.
First, few theories have explicitly addressed adolescence.
This life stage differs qualitatively from both childhood and adulthood [@bariolaChildAdolescentEmotion2011].
It is therefore questionable whether theories focused on different age groups can be generalized to adolescents.
<!--Relatedly, many relevant theories lack an explicit developmental component, which renders them and thus lack a well-substantiated
understanding of developmental processes[see
@bussTheoriesEmotionalDevelopment2019;
@croneUnderstandingAdolescencePeriod2012]. -->
Furthermore, few theories have comprehensively addressed important predictors of development in this life stage,
and none directly guide contemporary research in the field [@bussTheoriesEmotionalDevelopment2019].
Finally, existing theories vary widely in scope:
Some are broad and non-specific; others describe a specific phenomenon in detail,
but lack a broader perspective.
Broad theories can be used to frame a wide variety of research,
and specific theories are more suitable for deductive hypothesis generation.
To combine the strength of both, it would be beneficial to bridge these levels of analysis.
There is a need for integrative theory formation,
which could provide a unified framework to guide future empirical work.<!--[see @riedigerEmotionRegulationAdolescence2014].
Inductive analysis of the empirical literature may constitute a first step to this end.-->
<!--Such a theory should focus on -->
<!--
Barriola: What is most
* evident from the reviewed studies is that the samples
investigated have been mainly non-clinical, within the
early childhood period of development and have examined
predominantly maternal socialization factors.
* the vast
majority have been restricted to toddlers and young chil-
dren and have only examined the in?uence of maternal,
and not paternal, expression.
Most of the literature has addressed emotion regulation in adult or infant
samples. A smaller, but substantial, portion of the literature has examined
emotion regulation in adolescents.
-->

@@

<!--Whereas theory provides a top-down deductive understanding of adolescents' emotion regulation,-->
Literature reviews can shed a light on undertheorized relevant phenomena by
synthesizing inductive insights emerging from the empirical literature.
This is especially important given the noted absence of a single overarching theory [see @bussTheoriesEmotionalDevelopment2019].
One recent narrative review, in exploring which factors contribute to adolescents' emotion regulation development, clearly reflected the emphasis on neurocognitive factors and socialization that is evident from the theoretical literature [@riedigerEmotionRegulationAdolescence2014].
Other recent narrative reviews complement the aforementioned theoretical literature to a greater extent [@coe-odessEmergentEmotionsAdolescence2019; @bariolaChildAdolescentEmotion2011].
The right panel of Figure \@ref(fig:figbaseline) visualizes constructs uniquely covered by these narrative reviews,
relative to the preceding theoretical literature.
This illustrates in what respects these reviews reflect existing theory,
and in what respects they complement it.
<!--
Note that the left panel of Figure \@ref(fig:tmnetworks) visually summarizes the correlates of emotion regulation difficulties (labeled: *Dysregulation*) according to our theoretical review.
The left panel of Figure \@ref(fig:tmnetworks) summarizes key constructs covered by the theories reviewed above. 
In the subsequent text mining analysis, we will distinguish the topics covered in this theoretical review visually to highlight unique contributions of our approach.
This allows us to see in what respects the text mining analysis reflects existing theory,
and in what respects it complements it.-->

>>>>

---
title             : "Mapping Phenomena Relevant to Adolescent Emotion Dysregulation: A Text-mining Systematic Review"
shorttitle        : "ADOLESCENT EMOTION REGULATION"

@@

Adolescence is a developmentally sensitive period for emotion regulation [@zimmermannEmotionRegulationEarly2014]. 
<!--Adolescents experience biological, cognitive, and social changes that prompt new emotional
experiences and tax regulatory abilities [@steinbergAgeOpportunityLessons2014].
These coalescing changes temporarily restrict adolescents' capacity for emotion regulation 
[@zimmermannEmotionRegulationEarly2014],
resulting in more frequent and intense (negative) emotions [see @silkAdolescentsEmotionRegulation2003].-->
Although most adolescents successfully develop mature emotion regulation abilities,
as many as one in five develop psychopathology [@leeAdolescentMentalHealth2014]
with roots in emotion regulation difficulties
[@aldaoEmotionregulationStrategiesPsychopathology2010].
Given the prevalence of emotion regulation difficulties in adolescence,
their implications for mental health<!--and social functioning [@braetEmotionRegulationChildren2014]; @reindlSocializationEmotionRegulation2016]-->
and downstream cost to society,
it is important to have a comprehensive overview of the phenomena associated with adolescents' emotion regulation.
<!--Of particular interest are potential risk factors that render some youth susceptible to emotion regulation difficulties.-->
Despite the abundance of publications in this field,
it is difficult to obtain a comprehensive understanding of emotion regulation in adolescence [@riedigerEmotionRegulationAdolescence2014]:
Different (sub)disciplines have approached the topic in disparate ways [@riedigerEmotionRegulationAdolescence2014],
without consistent terminology [@bariolaChildAdolescentEmotion2011], conceptual frameworks [@stifterEmotionRegulation2019], or an overarching theoretical framework
[@bussTheoriesEmotionalDevelopment2019].
This has prompted calls for the integration of knowledge across research areas [@riedigerEmotionRegulationAdolescence2014] and consolidation of empirical research into overarching theory [e.g., @bussTheoriesEmotionalDevelopment2019].
Formal methods for theory construction state that the first step in this endeavor is to identify well-established relevant *empirical phenomena*,
defined as stable and general features of the world [@borsboomTheoryConstructionMethodology2020].
The current study used a text mining systematic review to provide a comprehensive map of the empirical phenomena relevant to adolescent emotion regulation.

Currently, few instruments exist for the identification of relevant phenomena; instead, theorists typically rely on expert opinion [@borsboomTheoryConstructionMethodology2020].
This paper argues that systematic reviews are a suitable and relatively more objective method for phenomena detection.
Traditional narrative reviews are limited, however, by small convenience samples,
confirmation bias,
and an undue emphasis on positive results [@littellEvidencebasedBiasedQuality2008].
These limitations are overcome by a novel method:
the text mining systematic review (TMSR).
Relative to narrative reviews, text mining can cover arbitrarily greater corpora,
and gleans insights from the literature by means of a relatively more transparent, objective, and reproducible process.
<!--Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008],
the inherent "fairness" of text mining gives emerging themes a chance to come to the fore.-->
This approach assumes the frequency with which phenomena recur in the literature to be indicative of their relevance.
If it is additionally assumed that the frequency with which phenomena are studied together is indicative of the extent to which they are thought to be associated,
then the co-occurrence of phenomena within publications can be interpreted as a rudimentary nomological network: a mapping of relationships between theoretically relevant phenomena [see @alaviAligningTheoryMethodology2018].
Such a network can serve as a starting point for the formulation of a proto-theory,
which would additionally involve abductive reasoning: specifying putative mechanisms to explain associations between the phenomena [@borsboomTheoryConstructionMethodology2020].
Text mining systematic reviews constitute an inductive and exploratory approach:
Their aim is phenomena detection based on generalization from observations (data) without appeal to theory [@haigExploratoryFactorAnalysis2005].
By identifying and mapping the phenomena relevant to adolescent emotion regulation,
such an inductive approach lays the groundwork for future theory development.

@@

p <- quote({
  # Set margins to 0
  par(mar=c(0.5,0.5,0.5,0.5),
      mfrow=c(1,2))
  plot(g1, edge.curved = 0, layout=l1,
       vertex.label.family = "sans",
       vertex.label.cex = 0.7,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "a")
  legend(x=-1.2, y=1.1, names(cat_cols), pch=legend_shapes, col=cat_cols, pt.bg=cat_cols, pt.cex=1.4, cex=.7, bty="n", ncol=1)
  plot(g2, edge.curved = 0, layout=l2,
       vertex.label.family = "sans",
       vertex.label.cex = 0.7,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "b")
})
saveRDS(g1, "bl_g1.RData")
saveRDS(g2, "bl_g2.RData")
saveRDS(l1, "bl_l1.RData")
saveRDS(l2, "bl_l2.RData")
# Save files

@@

```{r figbaseline, warning=FALSE, eval = TRUE, results='asis', out.width = "100%", fig.cap="Phenomena relevant to adolescents' emotion regulation according to theory (a) and narrative reviews (b; transparent nodes indicate constructs also present in the theory)."}
knitr::include_graphics("baseline_network.png") 
```
```{r allnetworks, warning=FALSE, eval = FALSE, results='asis', out.width = "100%", fig.cap="Correlates of emotion regulation difficulties according to theory (a), narrative reviews (b), and term co-occurrence in author keywords (c) and abstracts (d). Size of lines and nodes represents frequency. Dashed lines represent links not involving dysregulation. Transparent nodes indicate constructs also represented in the theory."}

@@

Although it has been argued that there is a paucity of theories *specific to* adolescent emotion regulation [see @bussTheoriesEmotionalDevelopment2019],
several *relevant* theories are commonly invoked in empirical work.
A brief review of the existing theoretical landscape is provided here
to contextualize the present systematic review
and assess to what extent phenomena identified through text mining mirror or complement existing theory.
Others have published detailed reviews of theories of emotional development [@bussTheoriesEmotionalDevelopment2019] and of empirical research on adolescent emotion regulation [@riedigerEmotionRegulationAdolescence2014].
The left panel of Figure \@ref(fig:figbaseline) visually summarizes the phenomena discussed in the theoretical literature reviewed below,
along with whether these were discussed as putative causes, outcomes, indicators, or protective factors in relation to emotion regulation.
This establishes a baseline for the original contributions of the text minning analysis.
<!-- 
Our text mining analysis of the phenomena empirical literature is best understood against a backdrop of -->

Two of the most general theories invoked to frame developmental research are the bioecological model [@bronfenbrennerBioecologicalModelHuman2007], and the transactional model [@sameroffUnifiedTheoryDevelopment2010].
The bioecological model describes how the `r keyw("environment")`
shapes individual development.
At the individual level, every person is imbued with `r keyw("biological")` predispositions,
and develops over time in interaction with contextual influences.
The most immediate source of contextual influences is the microsystem,
composed of primary caregivers and other close ties.
Other environmental influences originate in the macrosystem - including `r keyw("political")` and `r keyw("economic")` influences - and the exosystem,
consisting of `r keyw("cultural")` norms and `r keyw("values")`.
The transactional model [@sameroffUnifiedTheoryDevelopment2010] is compatible with the bioecological model,
but places a stronger emphasis on development as a product of reciprocal influences between child and environment.
The transactional model distinguishes between proximal influences, which roughly correspond to the microsystem in the bioecological model, and distal influences, which derive from
structural and environmental factors indirectly shaping development, like `r keyw("socio-economic status")`,
`r keyw("schools")`, and the `r keyw("community")` (macro- and exosystem).
With increasing age, these distal influences gain ground on proximal influences.
These two theories have a broader scope than most.
This means that they can be invoked to contextualize any developmental study,
but lack specificity - a shortcoming that curtails a theory's utility in generating hypotheses.
Domain-specific theories, by comparison, offer greater specificity.

One such domain-specific theory is Hall's notion of "storm and stress" in adolescence [see
@arnettAdolescentStormStress1999].
It describes how `r keyw("hormonal")` changes diminish `r keyw("self-control")` and increase `r keyw("reactivity")`, which in turn leads to emotion
regulation difficulties, increased `r keyw("conflict with parents")`, and `r keyw("risky behavior")`.
Contemporary theories recast this notion of diminished self-control and increased emotional reactivity as a normative change that facilitates emotional maturation at the risk of emotional
disturbance [@arnettAdolescentStormStress1999].
A potential limitation of conceptualizing emotion regulation difficulties as normative is that this underemphasizes heterogeneity in developmental trajectories [cf. @croneUnderstandingAdolescencePeriod2012].
<!--SH I am still not sure what the empirical contribution is - can you be more specific with it in this sentence? -->
<!--Several recent publications have undertaken the
monumental task of providing a comprehensive overview of theories of emotional
development [e.g., @bussTheoriesEmotionalDevelopment2019;
@holodynskiDevelopmentEmotionsEmotion2006;
@riedigerEmotionRegulationAdolescence2014], and of the empirical work
pertaining to emotions in adolescence
[@coe-odessEmergentEmotionsAdolescence2019].  -->

One theory of normative emotional development focuses on developmental influences to a greater extent [@sroufeEmotionalDevelopmentOrganization1995].
It posits that, as children grow older,
their increasing `r keyw("self-regulatory")` abilities
drive a transition from external emotion regulation by `r keyw("primary caregivers")`
toward autonomous emotion regulation.
This theory emphasizes `r keyw("social")` and `r keyw("cognitive influences")` as 
drivers of emotion regulation development.
Social influences mainly occur
through `r keyw("parental co-regulation")`, `r keyw("parenting behaviors")`, and parent-child
`r keyw("attachment")`.
Cognitive influences occur through the development of the `r keyw("central
nervous system")` (CNS), `r keyw("cognition")`, and `r keyw("self-regulation")`.
<!--In terms of level of analysis, Sroufe's and Hall's theories corresponds approximately to the individual- and microsystem level of the ecological model [@bronfenbrennerBioecologicalModelHuman2007].-->
This theory's relevance for adolescent research is somewhat diminished by its focus on early childhood.
However, research indicates that there is some continuity in developmental influences throughout childhood and adolescence; for instance, mothers' and fathers' unique roles in emotional development are similar in childhood and adolescence [see @vanlissaMothersFathersQuantitative2020; @vanlissaRoleFathersMothers2019].
The emphasis on socialization and `r keyw("neurocognitive development")`,
which is evident from both theories,
is the focus of several theories with a narrower scope.

One widely cited theory of socialization is the tripartite model [@morrisRoleFamilyContext2007].
It describes three pathways through which `r keyw("parents")` shape emotion regulation: `r keyw("modeling")`, `r keyw("parenting practices")`, and the
emotional `r keyw("family climate")`,
which subsumes attachment and `r keyw("marital relationship")` quality.
This theory acknowledges unique contributions by `r keyw("mothers")` and `r keyw("fathers")`, and the importance of siblings.
Others have adapted the tripartite model to describe peer socialization [@reindlSocializationEmotionRegulation2016], thus illustrating its wider usefulness.
A more abstract take on socialization is found in the internalization model of emotional development [@holodynskiDevelopmentEmotionsEmotion2006]. 
This theory describes the socialization of the role of emotion in `r keyw("communication")`, and the `r keyw("cultural")` symbolic function of emotion.

In research on neurocognitive aspects of adolescents' emotion regulation,
`r keyw("polyvagal theory")` has been influential [@porgesOrientingDefensiveWorld1995].
This theory examines `r keyw("emotional experience")` and -regulation in relation to autonomous nervous system functioning, respiratory sinus arrhythmia, and the stress response.
Although polyvagal theory is relevant for development,
it does not explicitly address it.
Among developmental neurocognitive theories,
there is some consensus that the developmental asymmetry between `r keyw("motivational")`-emotional and inhibitory `r keyw("brain")` circuits gives rise to a "maturity gap" in middle adolescence [@croneUnderstandingAdolescencePeriod2012].
According to the model of social-affective engagement and `r keyw("goal flexibility")` [@croneUnderstandingAdolescencePeriod2012],
adolescents' cognitive engagement is dynamically responsive to
social and motivational `r keyw("goal salience")`.
This flexibility <!--, on the one hand,-->
prepares adolescents to <!--effectively engage cognitive systems in novel
challenging situations in a way that facilitates--> develop mature regulatory
abilities, but also places them at risk of greater impulsivity - for example, in
pursuit of `r keyw("peer approval")`.
To a greater extent than related writings [cf. @craccoEmotionRegulationChildhood2017], this model focuses on adolescents diverging destinies;
why some youngsters flourish while others languish [@croneUnderstandingAdolescencePeriod2012].
This is important for understanding individual differences in development.
Although the relevance of this theory is enhanced by its focus on adolescence,
it only tangentially addresses emotion regulation.
Furthermore, whereas this theory addresses cognitive factors and peers as drivers of development, it underemphasizes other relevant phenomena, such as parenting [cf. @morrisRoleFamilyContext2007].
Micro-scale theories like these are useful in explaining specific phenomena in detail,
but do not provide a comprehensive understanding of adolescent's emotion regulation.
To this end, it is important to consider them in the context of a larger framework of relevant phenomena.

<!--Like Sroufe, Holodynski describes an age graded transition from
interpersonal to intrapersonal emotion regulation.
In doing so, Holodynski applies Vygotsky's theory of development to the domain of
emotion, and presents a detailed integrated model of emotional experience and
regulation.-->
Theories of emotion regulation offer further insight into
intra-individual drivers of emotion regulation development.
The influential process model describes emotion regulation throughout the progression from eliciting cue to ultimate response [@grossHandbookEmotionRegulation2013].
This model posits that individuals use `r keyw("strategies")` to modulate this process at different stages,
consciously or otherwise. 
Maladaptive emotion regulation strategies tend to result in more negative emotions,
diminished well-being, and greater strain in interpersonal relationships
[ @bellRelationshipsInputsOutputs2000].
However, comparative studies indicate that the adaptive versus maladaptive psychosocial consequences of specific strategies appear to be partly contingent on the `r keyw("cultural context")` [see
@bariolaChildAdolescentEmotion2011].
Similar to the process model, the "social
information processing" theory also describes the role of cognitive
processes and strategies in emotion regulation
[@lemeriseIntegratedModelEmotion2000].
One shortcoming of these theories for understanding adolescent emotion regulation 
is the lack of a developmental component.

@@

Despite the abundance of theory relevant to emotion regulation in adolescence,
the literature has several limitations.
First, few theories have explicitly addressed adolescence.
As this life stage differs qualitatively from both childhood and adulthood,
it is questionable whether theories focused on different age groups can be generalized to adolescents [@bariolaChildAdolescentEmotion2011].
<!--Relatedly, many relevant theories lack an explicit developmental component, which renders them and thus lack a well-substantiated
understanding of developmental processes[see
@bussTheoriesEmotionalDevelopment2019;
@croneUnderstandingAdolescencePeriod2012]. -->
Furthermore, few theories have comprehensively addressed important predictors of development in this life stage,
and none directly guide contemporary research in the field [@bussTheoriesEmotionalDevelopment2019].
Finally, existing theories vary widely in scope:
Some are broad and non-specific; others describe a specific phenomenon in detail,
but lack a broader perspective.
Broad theories can be used to frame a wide range of research,
and specific theories are more suitable for deductive hypothesis generation.
To combine the strength of both, it would be beneficial to bridge these levels of analysis.
There is a need for integrative theory formation,
which could provide a unified framework to guide future empirical work.<!--[see @riedigerEmotionRegulationAdolescence2014].
Inductive analysis of the empirical literature may constitute a first step to this end.-->
<!--Such a theory should focus on -->
<!--
Barriola: What is most
* evident from the reviewed studies is that the samples
investigated have been mainly non-clinical, within the
early childhood period of development and have examined
predominantly maternal socialization factors.
* the vast
majority have been restricted to toddlers and young chil-
dren and have only examined the in?uence of maternal,
and not paternal, expression.
Most of the literature has addressed emotion regulation in adult or infant
samples. A smaller, but substantial, portion of the literature has examined
emotion regulation in adolescents.
-->

@@

<!--Whereas theory provides a top-down deductive understanding of adolescents' emotion regulation,-->
Literature reviews can shed light on undertheorized relevant phenomena by
synthesizing inductive insights emerging from the empirical literature.
This is especially important given the noted absence of overarching theory [see @bussTheoriesEmotionalDevelopment2019].
One recent narrative review, in exploring which factors contribute to adolescents' emotion regulation development, clearly reflected the emphasis on neurocognitive factors and socialization that is evident from the theoretical literature [@riedigerEmotionRegulationAdolescence2014].
Other recent narrative reviews complement the theoretical literature to a greater extent by addressing relevant but undertheorized phenomena [@coe-odessEmergentEmotionsAdolescence2019; @bariolaChildAdolescentEmotion2011].
The right panel of Figure \@ref(fig:figbaseline) illustrates in what respects these narrative reviews reflect existing theory,
and in what respects they complement it.
<!--
Note that the left panel of Figure \@ref(fig:tmnetworks) visually summarizes the correlates of emotion regulation difficulties (labeled: *Dysregulation*) according to our theoretical review.
The left panel of Figure \@ref(fig:tmnetworks) summarizes key constructs covered by the theories reviewed above. 
In the subsequent text mining analysis, we will distinguish the topics covered in this theoretical review visually to highlight unique contributions of our approach.
This allows us to see in what respects the text mining analysis reflects existing theory,
and in what respects it complements it.-->

One seminal review discussed relevant phenomena at different levels of analysis [@bariolaChildAdolescentEmotion2011]. 
At the individual level, such relevant phenomena include temperament and biological factors, like neurocognitive development and genes.
At the level of proximal influences, the authors discuss socialization and modeling by parents, teachers, and peers.
Finally, at the level of distal influences, culture and the media are discussed.
This review also called for research on emotion regulation beyond early childhood and into adolescence, and on fathers' unique contributions -
a topic receiving increasing interest [see @pleckPaternalInvolvementRevised2004].
In line with these recommendations, several recent publications have investigated mothers' and fathers' unique roles in emotion regulation socialization from childhood to adolescence [see @vanlissaMothersFathersQuantitative2020] and throughout adolescence [see @vanlissaRoleFathersMothers2019].
